---
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

### *Taras Yaroshko, Liubomyr Oleksyuk, Yaroslav Klym*

## Introduction

During the past three weeks, you learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

All the terms on the right-hand side can be estimated from the data as
respective relative frequencies;\
see [this
site](https://monkeylearn.com/blog/practical-explanation-naive-bayes-classifier/)
for more detailed explanations.

## Data description

There are 5 datasets uploaded on the cms.

To determine your variant, take your team number from the list of teams
on cms and take *mod 5* - this is the number of your data set.

-   **0 - authors** This data set consists of citations of three famous
    writers: Edgar Alan Poe, Mary Wollstonecraft Shelley and HP
    Lovecraft. The task with this data set is to classify a piece of
    text with the author who was more likely to write it.

-   **1 - discrimination** This data set consists of tweets that have
    discriminatory (sexism or racism) messages or of tweets that are of
    neutral mood. The task is to determine whether a given tweet has
    discriminatory mood or does not.

-   **2 - fake news** This data set contains data of American news: a
    headline and an abstract of the article. Each piece of news is
    classified as fake or credible. The task is to classify the news
    from test.csv as credible or fake.

-   **3 - sentiment** All the text messages contained in this data set
    are labeled with three sentiments: positive, neutral or negative.
    The task is to classify some text message as the one of positive
    mood, negative or neutral.

-   **4 - spam** This last data set contains SMS messages classified as
    spam or non-spam (ham in the data set). The task is to determine
    whether a given message is spam or non-spam.

Each data set consists of two files: *train.csv* and *test.csv*. The
first one you will need find the probabilities distributions for each of
the features, while the second one is needed for checking how well your
classifier works.

```{r}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed
library(tidytext)
library(readr)
library(dplyr)
library(ggplot2)
```

## Instructions

-   The first step is data pre-processing, which includes removing
    punctuation marks and stop words

-   represent each message as a bag-of-words

-   using the training set, calculate all the conditional probabilities
    in formula (1)

-   use those to predict classes for messages in the test set

-   evaluate effectiveness of the classifier by calculating the
    corresponding metrics

-   shortly summarize your work

-   do not forget to submit both the (compiled) Rmd source file and the
    .html output

### Data pre-processing

-   Read the *.csv* data files.
-   Сlear your data from punctuation or other unneeded symbols.
-   Clear you data from stop words. You don't want words as is, and, or
    etc. to affect your probabilities distributions, so it is a wise
    decision to get rid of them. Find list of stop words in the cms
    under the lab task.
-   Represent each test message as its bag-of-words. Here:
    <https://machinelearningmastery.com/gentle-introduction-bag-words-model/>
    you can find general introduction to the bag-of-words model and
    examples on to create it.
-   It is highly recommended to get familiar with R dataframes, it would
    make the work much easier to do.
-   Useful links:
    -   <https://steviep42.github.io/webscraping/book/bagofwords.html#tidytext> -
        example of using *tidytext* to count frequencies of the words.
    -   Basics of Text Mining in R:
        <http://rstudio-pubs-static.s3.amazonaws.com/256588_57b585da6c054349825cba46685d8464.html>
        . Note that it also includes an example on how to create a bag
        of words from your text document.

```{r}
list.files(getwd())
list.files("data/0-authors")
```

```{r}
test_path <- "data/0-authors/test.csv"
train_path <- "data/0-authors/train.csv"

stop_words <- read_file("stop_words.txt")
# https://stackoverflow.com/questions/27195912/why-does-strsplit-return-a-list
splitted_stop_words <- strsplit(stop_words, split='\r\n')
splitted_stop_words <- splitted_stop_words[[1]]
```

```{r}
train <-  read.csv(file = train_path, stringsAsFactors = FALSE)
test <-  read.csv(file = test_path, stringsAsFactors = FALSE)
```

```{r}
# note the power functional features of R bring us! 
tidy_text <- unnest_tokens(train, 'splitted', 'text', token="words") %>%
             filter(!splitted %in% splitted_stop_words)
```

```{r}
authors_frame <- data.frame("word", "HP Lovecraft", "Edgar Alan Poe", "Mary Wollstonecraft Shelley ")

```

### Data visualization

Each time you work with some data, you need to understand it before you
start processing it. R has very powerful tools to make nice plots and
visualization. Show what are the most common words for negative and
positive examples as a histogram, word cloud etc. Be creative!

```{r}
sentences_by_author <- tidy_text %>%

  dplyr::group_by(author) %>%

  dplyr::summarise(n = n())

ggplot(data = sentences_by_author, mapping = aes(x = author, y = n, fill = author)) +
  geom_col(show.legend = F) + ylab(label = "number of sentences") + ggtitle("Number of sentences per author")
```

## Classifier implementation

```{r}
naiveBayes <- setRefClass("naiveBayes",
                          
       # here it would be wise to have some vars to store intermediate result
       # frequency dict etc. Though pay attention to bag of wards! 
       fields = list(author_df = "data.frame"),
       methods = list(
                    # prepare your training data as X - bag of words for each of your
                    # messages and corresponding label for the message encoded as 0 or 1 
                    # (binary classification task)
                    fit = function(X)
                    {
                         # P(some_word|author) = (amount_of_some_word_in_author + 1)/(all_word_in_author + all_diff_word)
                      tidy_text <- unnest_tokens(X, 'splitted', 'text', token = "words") %>% 
                        filter(!splitted %in% splitted_stop_words)
                      all_diff_word <- tidy_text %>% count(splitted,sort=TRUE) %>% nrow()
                      words_by_authors <- tidy_text %>% count(splitted,sort=TRUE, author)
                      all_word_in_edgar <- sum(words_by_authors$n[which(words_by_authors$author=="Edgar Alan Poe")])
                      all_word_in_lovecraft <- sum(words_by_authors$n[which(words_by_authors$author=="HP Lovecraft")])
                      all_word_in_mary <- sum(words_by_authors$n[which(words_by_authors$author=="Mary Wollstonecraft Shelley ")])
                      for (word in unique(tidy_text$splitted)){
                        amount_of_some_word_in_edgar <- words_by_authors$n[which((words_by_authors$splitted==word) & (words_by_authors$author=="Edgar Alan Poe"))]
                        if (length(amount_of_some_word_in_edgar) == 0){
                          amount_of_some_word_in_edgar <- 0
                        }
                        prob_word_in_edgar = (amount_of_some_word_in_edgar + 1)/(all_word_in_edgar + all_diff_word)
                        author_df[word, "Edgar Alan Poe"] <<- prob_word_in_edgar
                        amount_of_some_word_in_lovecraft <- words_by_authors$n[which((words_by_authors$splitted==word) & (words_by_authors$author=="HP Lovecraft"))]
                        if (length(amount_of_some_word_in_lovecraft) == 0){
                          amount_of_some_word_in_lovecraft <- 0
                        }
                        prob_word_in_lovecraft = (amount_of_some_word_in_lovecraft + 1)/(all_word_in_lovecraft + all_diff_word)
                        author_df[word, "HP Lovecraft"] <<- prob_word_in_lovecraft
                        
                        amount_of_some_word_in_mary <- words_by_authors$n[which((words_by_authors$splitted==word) & (words_by_authors$author=="Mary Wollstonecraft Shelley "))]
                        if (length(amount_of_some_word_in_mary) == 0){
                          amount_of_some_word_in_mary <- 0
                        }
                        prob_word_in_mary = (amount_of_some_word_in_mary + 1)/(all_word_in_mary + all_diff_word)
                        author_df[word, "Mary Wollstonecraft Shelley "] <<- prob_word_in_mary
                      }
                    
                    },
                    
                    # return prediction for a single message 
                    predict = function(message)
                    {
                      splitted_message <- strsplit(message, split=' ')
                      splitted_message <- splitted_message[[1]]
                      prob_of_edgar <- nrow(train%>%filter(author == "Edgar Alan Poe"))/nrow(train)
                      prob_of_lovecraft <- nrow(train%>%filter(author == "HP Lovecraft"))/nrow(train)
                      prob_of_mary <- nrow(train%>%filter(author == "Mary Wollstonecraft Shelley "))/nrow(train)
                      
                      prob_massage_in_edgar = 1
                      prob_massage_in_lovecraft = 1
                      prob_massage_in_mary = 1
                      
                      for (word in splitted_message) {
                        if (!(is.na(author_df[word, "Edgar Alan Poe"]))){
                          prob_massage_in_edgar = prob_massage_in_edgar *
                            author_df[word, "Edgar Alan Poe"]
                        }
                        
                        if (!(is.na(author_df[word, "HP Lovecraft"]))){
                          prob_massage_in_lovecraft = prob_massage_in_lovecraft * author_df[word, "HP Lovecraft"]
                        }
                        if (!(is.na(author_df[word, "Mary Wollstonecraft Shelley "])))
                            {
                          prob_massage_in_mary = prob_massage_in_mary * author_df[word, "Mary Wollstonecraft Shelley "]
                        }
                      }
                      
                      prob_massage_in_edgar = prob_massage_in_edgar * prob_of_edgar
                      prob_massage_in_lovecraft = prob_massage_in_lovecraft * prob_of_lovecraft
                      prob_massage_in_mary = prob_massage_in_mary * prob_of_mary
                      max_prob <- max(prob_massage_in_edgar, prob_massage_in_lovecraft, prob_massage_in_mary)
                      if (max_prob == prob_massage_in_edgar){
                        return("Edgar Alan Poe")
                      }
                      else if (max_prob == prob_massage_in_lovecraft){
                        return("HP Lovecraft")
                      }
                      else{
                        return("Mary Wollstonecraft Shelley ")
                      }
                    },
                    score = function(X_test)
                    {
                      correct <- 0
                      for (i in 1:nrow(X_test)){
                        predicted <- model$predict(X_test[i, "text"])
                        if (X_test$author[i] == predicted){
                          correct <- correct + 1
                        }
                      }
                      return(correct / nrow(X_test))
                    }
))

model = naiveBayes(author_df = data.frame())
model$fit(train)
```

```{r}
# model$author_df
model$predict("Lying on its side for it could not stand up alone was the exotic spiky figure which in his monstrous dream he had broken off the fantastic balustrade.")
model$score(test)
```

## Measure effectiveness of your classifier

-   Accuracy of our classifier is 0.7854954.

## Conclusions

Here we implemented 3 methods: fit, predict and score.

Fit creates dataframe with words as indexes of raw and authors as
indexes of column. Values are probabilities of getting this word with
condition that we already know an author. We find it using a formula :

P(some_word\|author) = (amount_of_some_word_in_author +
1)/(all_word_in_author + all_diff_word)

Predict finds which author is most likely wrote given message. Here we
use this formula for every author and return author with the biggest
value of probability:

P(author\|message) = P(some_word1\|author) \* P(some_word2\|author) ...
\* P(some_word{len_of_message}\|author)  \* P(author)

Score checks whether predict was correct and count percent of correct
answers

The biggest plus of this method are that it is easy in implementation.

Cons are that it just searches by words, and not phrases, limited by
words, that we had in train, there better ways to predict.
